---
title: GenomicSEM of MDD symptoms
author: Mark Adams, Bradley Jermy, Jackson Thorp, Andrew Grotzinger, Michel Nivard 
output:
  html_document:
    toc: TRUE
    code_folding: "show"
    number_sections: TRUE
    df_print: kable
  md_document:
    variant: markdown_github
---

Conduct SEM model of MDD symptom genomic correlation matrix

```{r results='hide'}

library(GenomicSEM)
library(corrplot)

```

Read in LDSC structure for UKB CIDI symptoms
```{r}

cidi_covstruct <- dget('ldsc/ukb_mhq_cidi_covstruct.deparse.R')

# the +/- in weight change column names doesn't play well with
# lavaan model syntax
dimnames(cidi_covstruct$S)[[2]][10:11] <- c('WeightInc', 'WeightDec')

```

Plot correlation matrix (truncate correlations > 1)

```{r cidi_covstruct_corr}

cidi_cor <- cov2cor(cidi_covstruct$S)
cidi_cor[which(cidi_cor > 1)] <- 1
rownames(cidi_cor) <- colnames(cidi_cor)

corrplot(cidi_cor, 'square')

```

Fit common factor model (all symptoms)

```{r}

gfactor_fit <- commonfactor(cidi_covstruct)

```

```{r}

gfactor_fit$results

gfactor_fit$modelfit

```

Common factor model (excluding directional symptoms that have zero or negative genetic correlations with each other)

```{r}

symptomfactor_model <- "F1 =~ NA*Mood + Anhed + Cnctr + Death + Sleep + Tired + Worth + Weight
F1 ~~ 1*F1"

symptomfactor_fit <- usermodel(cidi_covstruct, model=symptomfactor_model)

```

```{r}

symptomfactor_fit$results

symptomfactor_fit$modelfit

```

Common factor model with directional symptoms but excluding their non-directional parent symptoms

```{r}

valencefactor_model <- "F1 =~ NA*Mood + Anhed + Cnctr + Death + Insomn + Hypersom + Tired + Worth + WeightInc + WeightDec
F1 ~~ 1*F1"

valencefactor_fit <- usermodel(cidi_covstruct, model=valencefactor_model)

```

```{r}

valencefactor_fit$results

valencefactor_fit$modelfit

```

Fit model close to Kendler…Neale (no psychomotor symptom).

```{r}

kendler...neale_model <- "
F1 =~ NA*Mood + Anhed + Worth
F2 =~ NA*Cnctr + Death + Worth
F3 =~ NA*Sleep + Tired + Weight
F1 ~~ 1*F1
F2 ~~ 1*F2
F3 ~~ 1*F3
F1 ~~ F2 + F3
F2 ~~ F3"
kendler...neale_fit <- usermodel(cidi_covstruct, model=kendler...neale_model)

```

Add constraints to the model to limit factor correlations to < 1 and residual variances to > 0.

```{r}

kendler...neale_constrain_model <- "
F1 =~ NA*Mood + Anhed + Worth
F2 =~ NA*Cnctr + Death + Worth
F3 =~ NA*Sleep + Tired + Weight
F1 ~~ 1*F1
F2 ~~ 1*F2
F3 ~~ 1*F3
F1 ~~ F2 + F3
F2 ~~ b*F3
Anhed~~a*Anhed
a > .001
b < 1"

kendler...neale_constrain_fit <- usermodel(cidi_covstruct, estimation = "DWLS", model=kendler...neale_constrain_model)

```

```{r}

kendler...neale_constrain_fit$modelfit

kendler...neale_constrain_fit$results

```
