---
title: GenomicSEM of MDD symptoms
author: Mark Adams, Bradley Jermy, Jackson Thorp, Andrew Grotzinger, Michel Nivard 
output:
  github_document
---


# Setup

## R packages

R version

```{r version}

R.version

```

Package installation

```{r packages, output='hide', warning=FALSE, message=FALSE, eval=FALSE}

required_packages <- c('devtools', 'readr', 'tidyr', 'dplyr', 'ggplot2', 'stringr', 'corrplot')
for(pack in required_packages) if(!require(pack, character.only=TRUE)) install.packages(pack)

if(!require(GenomicSEM)) remotes::install_github("MichelNivard/GenomicSEM")

```

GenomicSEM version

```{r requires, warning=FALSE, messages=FALSE}

require(readr)
require(tidyr)
require(stringr)
require(dplyr)
require(ggplot2)
require(corrplot)
require(GenomicSEM)

packageVersion("GenomicSEM")

```

# Symptom labels

MDD DSM symptoms are numbered 1-9:

```{r symptom_labels}

# plot labels

dsm_mdd_symptoms_labels <-
read_delim("
MDD1;Mood;Mood;Dep
MDD2;Interest;Interest;Anh
MDD3;Weight⇅;Weight⇆;App
MDD3a;Weight⇊;Weight⇇;AppDec
MDD3b;Weight⇈;Weight⇉;AppInc
MDD4;Sleep⇅;Sleep⇆;Sle
MDD4a;Sleep⇊;Sleep⇇;SleDec
MDD4b;Sleep⇈;Sleep⇉;SleInc
MDD5;Motor⇅;Motor⇆;Moto
MDD5a;Motor⇈;Motor⇉;MotoInc
MDD5b;Motor⇊;Motor⇇;MotoDec
MDD6;Fatigue;Fatigue;Fatig
MDD7;Guilt;Guilt;Guilt
MDD8;Concentrate;Concentrate;Conc
MDD9;Suicidality;Suicidality;Sui
", col_names=c('ref', 'h', 'v', 'abbv'), delim=';')

dsm_mdd_symptoms_reference <-
read_delim("
MDD1;Depressed mood most of the day, nearly every day
MDD2;Markedly diminished interest or pleasure in all, or almost all, activities most of the day, nearly every day
MDD3;Significant change in weight or appetite
MDD3a;Significant weight loss or decrease in appetite
MDD3b;Significant weight gain or increase in appetite
MDD4;Sleeping too much or not sleeping enough
MDD4a;Insomnia nearly every day
MDD4b;Hypersomnia nearly every day
MDD5;Changes in speed/amount of moving or speaking
MDD5a;Psychomotor agitation nearly every day
MDD5b;Psychomotor slowing nearly every day
MDD6;Fatigue or loss of energy nearly every day
MDD7;Feelings of worthlessness or excessive or inappropriate guilt
MDD8;Diminished ability to think or concentrate, or indecisiveness
MDD9;Recurrent thoughts of death or suicide or a suicide attempt or a specific plan for attempting suicide
", col_names=c('Reference', 'Description'), delim=';')

dsm_mdd_symptoms_reference %>%
left_join(dsm_mdd_symptoms_labels, by=c('Reference'='ref')) %>%
select(Reference, Abbreviation=abbv, Label=h, Description)

```

# GenomicSEM covariance structure

```{r covstruct}

covstruct_prefix <- 'clin.comm.covstruct'
covstruct_r <- file.path('ldsc', paste(covstruct_prefix, 'deparse.R', sep='.'))
covstruct_rds <- file.path('ldsc', paste(covstruct_prefix, 'rds', sep='.'))

symptoms_covstruct <- dget(covstruct_r)

sumstats_prevs <- read_tsv(file.path('ldsc', paste(covstruct_prefix, 'prevs', 'txt', sep='.')))

```

Rename samples: AGDS/GS/PGC is the Case-only **Clin**incal meta-analysis sample (`Clin`) and ALSPAC/EstBB/UKB is the **Comm**unity meta-analysis sample (`Comm`); and rename symptoms numbers (`MDD1`, `MDD2`) to abbreviations (`Dep`, `Anh`). There are also extra measures of `MDD1` and `MDD2` from **UKB** Baseline data. 

```{r sem_labels}

cohorts_sample_symptoms <-
sumstats_prevs %>%
left_join(dsm_mdd_symptoms_labels, by=c('symptom'='ref')) %>%
select(cohorts, symptom, trait_name, abbv) %>%
mutate(Sample=case_when(cohorts %in% 'Clin' ~ 'Clin',
                        cohorts %in% 'Comm' ~ 'Comm',
                        cohorts %in% 'UKBt' ~ 'Ukb',
                        TRUE ~ NA_character_)) %>%
mutate(sample_symptom=paste0(Sample, abbv))

sample_symptoms <- cohorts_sample_symptoms$sample_symptom
names(sample_symptoms) <- cohorts_sample_symptoms$trait_name

# rename traits in covstruct
dimnames(symptoms_covstruct$S)[[2]] <-
as.vector(sample_symptoms[dimnames(symptoms_covstruct$S)[[2]]])

```

# Structural models

Symptoms with positive variances

```{r}

symptoms_S_var <- diag(symptoms_covstruct$S)
names(symptoms_S_var) <- dimnames(symptoms_covstruct$S)[[2]]

symptoms_S_var[which(symptoms_S_var > 0)]

```

## Common factor

Common factor across symptoms from both cohorts, as a general MDD factor

```{r commonfactor, cache=TRUE}

commonfactor.model <- "
MDD =~ NA*ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec + ClinGuilt + ClinSui + CommDep + CommAnh + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig + CommGuilt + CommConc + CommSui + UkbDep + UkbAnh

MDD ~~ 1*MDD
"

commonfactor.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=commonfactor.model, imp_cov=TRUE)

commonfactor.fit$modelfit
commonfactor.fit$results[c(1,2,3,6,7)] %>%
     filter(lhs == 'MDD') %>%
     mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
     print(digits=2)

```

## Directional symptoms

Common factor with residual correlations among paired directional symptoms.

```{r common_dir, cache=TRUE}

common_dir.model <- "
MDD =~ NA*ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec + ClinGuilt + ClinSui + CommDep + CommAnh + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig + CommGuilt + CommConc + CommSui + UkbDep + UkbAnh

ClinAppDec ~~ ClinAppInc
ClinSleDec ~~ ClinSleInc
ClinMotoInc ~~ ClinMotoDec
CommAppDec ~~ CommAppInc
CommSleDec ~~ CommSleInc

MDD ~~ 1*MDD
"

common_dir.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=common_dir.model, imp_cov=TRUE)

common_dir.fit$modelfit
common_dir.fit$results[c(1,2,3,6,7)] %>%
     filter(lhs == 'MDD' | (lhs != 'MDD' & lhs != rhs)) %>%
     mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
     print(digits=2)

```

## Ascertainment-specific factors

### Clinical-Community

Symptoms were assessed in Clinical and Community samples. Make factors representing sample type.

```{r clin_comm, cache=TRUE}

clin_comm.model <- "
CLIN =~ NA*ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec + ClinGuilt + ClinSui
COMM =~ NA*CommDep + CommAnh + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig + CommGuilt + CommConc + CommSui + UkbDep + UkbAnh

CLIN ~~ 1*CLIN
COMM ~~ 1*COMM

"

clin_comm.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=clin_comm.model, imp_cov=TRUE)

clin_comm.fit$modelfit
clin_comm.fit$results[c(1,2,3,6,7,9)] %>%
     filter(lhs %in% c('CLIN', 'COMM'), rhs %in% c('CLIN', 'COMM'), lhs != rhs) %>%
     mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
     print(digits=2)

```

### Gating symptoms

Then consider the community cardinal symptoms (Depression and Anhedonia), which are gating items to the rest of the community-sample symptoms.

```{r gate, cache=TRUE}

gate.model <- "
MDD =~ NA*ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec + ClinGuilt + ClinSui + CommDep + CommAnh + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig + CommGuilt + CommConc + CommSui + UkbDep + UkbAnh
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

MDD ~~ 1*MDD
GATE ~~ 1*GATE
MDD ~~ 0*GATE
"

gate.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=gate.model, imp_cov=TRUE)

gate.fit$modelfit
gate.fit$results[c(1,2,3,6,7)] %>%
     filter(lhs %in% c('GATE')) %>%
     mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
     print(digits=2)

```

### Gate-Community-Clinical (Spectrum)

Comparison of ascertainment and measurement. Gating symptoms from community sample (distinguish controls from subthreshold), symptoms from community sample symptoms (subthreshold from cases), and clinical cohort symptoms (distinguish cases from each other)

```{r measure, cache=TRUE}

measure.model <- "
CLIN =~ NA*ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec + ClinGuilt + ClinSui
COMM =~ NA*CommDep + CommAnh + UkbDep + UkbAnh + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig + CommGuilt + CommConc + CommSui
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

CLIN ~~ 1*CLIN
GATE ~~ 1*GATE
COMM ~~ 1*COMM
GATE ~~ 0*COMM + 0*CLIN
"

measure.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=measure.model, imp_cov=TRUE)

measure.fit$modelfit
measure.fit$results[c(1,2,3,6,7, 9)] %>%
     filter(lhs %in% c('CLIN', 'COMM', 'GATE'), rhs %in% c('CLIN', 'COMM', 'GATE'), lhs != rhs) %>%
     mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
     print(digits=2)

```

Pathway from Gating factor to Community factor

```{r measure_dep, eval=FALSE, cache=TRUE}

measure_dep.model <- "
CLIN =~ NA*ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec + ClinGuilt + ClinSui
CIDI =~ NA*CommDep + CommAnh + UkbDep + UkbAnh + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig + CommGuilt + CommConc + CommSui
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh
CIDI ~ GATE

CLIN ~~ 1*CLIN
GATE ~~ 1*GATE

uk1 > 0.001
UkbDep ~~ uk1*UkbDep
"

measure_dep.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=measure_dep.model, imp_cov=TRUE)

measure_dep.fit$modelfit
measure_dep.fit$results[c(1,2,3,6,7, 9)] %>%
	 filter(lhs %in% c('CLIN', 'COMM', 'GATE'), rhs %in% c('CLIN', 'COMM', 'GATE'), lhs != rhs) %>%
	 mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
	 print(digits=2)

```

### MDD Subtypes

Group symptoms from the clinical cohorts (that had positive genetic variance) together with the same symptoms from the community cohorts (to create a dimension of MDD subtypes), then a separate factor for all other community cohort symptoms (that separates cases from controls) 

```{r subtype, cache=TRUE}

subtype.model <- "
SUBTYPE =~ NA*ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec + ClinGuilt + ClinSui + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommSui
MDD =~ NA*CommDep + CommAnh + UkbDep + UkbAnh + CommFatig + CommGuilt + CommConc 

SUBTYPE ~~ 1*SUBTYPE
MDD ~~ 1*MDD
"

subtype.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=subtype.model, imp_cov=TRUE)

subtype.fit$modelfit
subtype.fit$results[c(1,2,3,6,7)] %>%
     filter(lhs %in% c('SUBTYPE', 'MDD'), rhs %in% c('SUBTYPE', 'MDD'), lhs != rhs) %>%
     mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
     print(digits=2)

```

## Two-factor models

[Elhai Psychiat Res 2012](https://www.sciencedirect.com/science/article/pii/S0165178112002685) compared 3 two-factor models

### Psychological-Somatic (Elhai Model 2a)

[Krause Rehab Psychol 2008](https://psycnet.apa.org/record/2008-17022-011), [Krause Arch Psys Med Rehab 2010](https://www.sciencedirect.com/science/article/pii/S0003999310002443):

> the 2-factor solution with 3 somatic items (sleep disturbance, poor energy, appetite change) was a better solution than either a unidimensional model or 2-factor model that included psychomotor slowing as a fourth somatic item


```{r psych_soma, cache=TRUE}

psych_soma.model <- "
PSYCH =~ NA*ClinMotoInc + ClinMotoDec + ClinGuilt + ClinSui + CommDep + CommAnh + CommGuilt + CommConc + CommSui + UkbDep + UkbAnh
SOMA =~ NA*ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig

GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

PSYCH ~~ 1*PSYCH
SOMA ~~ 1*SOMA
GATE ~~ 1*GATE
GATE ~~ 0*PSYCH + 0*SOMA
"

psych_soma.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=psych_soma.model, imp_cov=TRUE)

psych_soma.fit$modelfit
psych_soma.fit$results[c(1,2,3,6,7)] %>%
     filter(lhs %in% c('PSYCH', 'SOMA'), rhs %in% c('PSYCH', 'SOMA'), lhs != rhs) %>%
     mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
     print(digits=2)

```

### Psychological-Vegetative (Elhai Model 2b)

```{r psych_veg, cache=TRUE}

psych_veg.model <- "
PSYCH =~ NA*ClinGuilt + ClinSui + CommDep + CommAnh + CommGuilt + CommConc + CommSui + UkbDep + UkbAnh
NEUROVEG =~ NA*ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig

GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

PSYCH ~~ 1*PSYCH
NEUROVEG ~~ 1*NEUROVEG
GATE ~~ 1*GATE
GATE ~~ 0*PSYCH + 0*NEUROVEG
"
psych_veg.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=psych_veg.model, imp_cov=TRUE)

psych_veg.fit$modelfit
psych_veg.fit$results[c(1,2,3,6,7)] %>%
     filter(lhs %in% c('PSYCH', 'NEUROEG'), rhs %in% c('PSYCH', 'NEUROEG'), lhs != rhs) %>%
     mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
     print(digits=2)

```


### Affective-Neurovegetative (Elhai Model 2c)

```{r affect_neuroveg, cache=TRUE}

affect_neuroveg.model <- "
AFFECT =~ NA*ClinGuilt + ClinSui + CommDep + CommAnh + CommGuilt + CommSui + UkbDep + UkbAnh
NEUROVEG =~ NA*ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig + CommConc

GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

AFFECT ~~ 1*AFFECT
NEUROVEG ~~ 1*NEUROVEG
GATE ~~ 1*GATE
GATE ~~ 0*AFFECT + 0*NEUROVEG
"
affect_neuroveg.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=affect_neuroveg.model, imp_cov=TRUE)

affect_neuroveg.fit$modelfit
affect_neuroveg.fit$results[c(1,2,3,6,7)] %>%
     filter(lhs %in% c('AFFECT', 'NEUROVEG'), rhs %in% c('AFFECT', 'NEUROVEG'), lhs != rhs) %>%
     mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
     print(digits=2)

```


## Three factor models

### Cognitive-Mood-Neuroveg (Kendler Neale) model

```{r cog_mood_neuroveg, cache=TRUE}

cog_mood_neuroveg.model <- "
COG =~ NA*ClinGuilt + ClinSui + CommGuilt + CommConc + CommSui
MOOD =~ NA*ClinGuilt + CommDep + CommAnh + CommGuilt + UkbDep + UkbAnh
NEUROVEG =~ NA*ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig

GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

COG ~~ 1*COG
MOOD ~~ 1*MOOD
NEUROVEG ~~ 1*NEUROVEG
GATE ~~ 1*GATE
GATE ~~ 0*COG + 0*MOOD + 0*NEUROVEG
"
cog_mood_neuroveg.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=cog_mood_neuroveg.model, imp_cov=TRUE)

cog_mood_neuroveg.fit$modelfit
cog_mood_neuroveg.fit$results[c(1,2,3,6,7)] %>%
     filter(lhs %in% c('COG', 'MOOD', 'NEUROVEG'), rhs %in% c('COG', 'MOOD', 'NEUROVEG'), lhs != rhs) %>%
     mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
     print(digits=2)

```

### Cognitive-Appetite-Vegetative (van Loo)


```{r cog_app_veg, cache=TRUE}

cog_app_veg.model <- "
COGMOOD =~ NA*CommDep + CommAnh + UkbDep + UkbAnh + ClinSleDec + CommSleDec + ClinMotoInc + ClinGuilt + CommGuilt + ClinSui + CommSui
APP =~ NA*CommAppDec + CommAppInc + ClinAppInc + ClinAppDec
LETH =~ NA*ClinSleInc + CommSleInc + ClinMotoDec + CommFatig + CommConc
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

COGMOOD ~~ 1*COGMOOD
APP ~~ 1*APP
LETH ~~ 1*LETH
GATE ~~ 1*GATE
GATE ~~ 0*COGMOOD + 0*APP + 0*LETH
"
cog_app_veg.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=cog_app_veg.model, imp_cov=TRUE)

cog_app_veg.fit$modelfit
cog_app_veg.fit$results[c(1,2,3,6,7, 9)] %>%
	 filter(lhs %in% c('COGMOOD', 'APP', 'LETH'), rhs %in% c('COGMOOD', 'APP', 'LETH'), lhs != rhs) %>%
	 mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
	 print(digits=2)

```

### Melancholic and atypical

Account for directionality of symptoms using melancholic and atypical classifications, plus remaining affective/cognitive symptoms

```{r mel_aty_afc, cache=TRUE}

mel_aty_afc.model <- "
MEL =~ NA*CommAnh + UkbAnh + ClinSleDec + CommSleDec + ClinMotoInc + ClinAppDec + CommAppDec + ClinGuilt + CommGuilt
ATY =~ NA*ClinAppInc + CommAppInc + ClinSleInc + ClinMotoDec + CommSleInc + CommFatig
AFC =~ NA*CommDep + UkbDep + ClinSui + CommConc + CommSui 
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

MEL ~~ 1*MEL
ATY ~~ 1*ATY
AFC ~~ 1*AFC
GATE ~~ 1*GATE
GATE ~~ 0*AFC + 0*MEL + 0*ATY

mat < 0.99
maf < 0.99
MEL ~~ mat*ATY
MEL ~~ maf*AFC
"

mel_aty_afc.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=mel_aty_afc.model, imp_cov=TRUE)

mel_aty_afc.fit$modelfit
mel_aty_afc.fit$results[c(1,2,3,6,7,9)] %>%
	 filter(rhs %in% c('MEL', 'ATY', 'AFC'), lhs != rhs) %>%
	 mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
	 print(digits=2)

```

### Multivariate GWAS model

Symptoms well-powered enough for multivariate GWAS meta-analysis

```{r gwas}
gwas_meta.model <- "
META =~ NA*CommDep + CommSleDec + CommFatig + CommGuilt + CommConc + CommSui + UkbDep

META ~~ 1*META
"

gwas_meta.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=gwas_meta.model, imp_cov=TRUE)

gwas_meta.fit$modelfit
gwas_meta.fit$results[c(1,2,3,6,7)] %>%
     filter(lhs == 'META') %>%
     mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
     print(digits=2)
```

### Model comparisons

```{r model_list}

model_list=
list("A"=list(abbrev="Depr", name="Depression", model=commonfactor.fit),
     "B"=list(abbrev = "Case-Comm", name="Case-Community", model=clin_comm.fit),
     "C"=list(abbrev="Depr-Gate", name="Depression-Gating", model=gate.fit),
     "D"=list(abbrev="Case-Comm-Gate", name="Case-Community-Gating", model=measure.fit),
     "E"=list(abbrev = "Psyc-Soma", name="Psychological-Somatic", model=psych_soma.fit),
     "F"=list(abbrev = "Psyc-Neuv", name="Psychological-Neurovegetative", model=psych_veg.fit),
     "G"=list(abbrev = "Affc-Neuv", name="Affective-Neurovegetative", model=affect_neuroveg.fit),
     "H"=list(abbrev = "Cog-Mood-Neuv", name="Cognitive-Mood-Neurovegetative", model=cog_mood_neuroveg.fit),
	 "I"=list(abbrev = "CogMood-App-Leth", name="Cognitive/Mood-Appetite-Lethargy", model=cog_app_veg.fit),
	 "J"=list(abbrev = "AffCog-Melc-Atyp", name="Affective/Cognitive-Melancholic-Atypical", model=mel_aty_afc.fit)
     )

model_fits <- 
data.frame(Model=sapply(model_list, function(m) m$abbrev),
           Name=sapply(model_list, function(m) m$name)) %>%
bind_cols(
bind_rows(
lapply(model_list, function(m) m$model$modelfit)
))
rownames(model_fits) <- NULL

knitr::kable(
model_fits %>%
mutate(dAIC=AIC-min(AIC)) %>%
mutate_if(is.numeric, ~signif(., 4))
)
```

The Mood-Appetite-Vegetative model is the best, SRMR is high across all the models, indicating that there are high residual correlations. 

## Modifications


### Combined factors

Combine Mood/Cognitive/Vegetative symptoms into a single factor

```{r cogmoodleth_app, cache=TRUE}

cogmoodleth_app.model <- "
COGMOODLETH =~ NA*CommDep + CommAnh + UkbDep + UkbAnh + ClinSleDec + CommSleDec + ClinMotoInc + ClinGuilt + CommGuilt + ClinSui + CommSui + ClinSleInc + CommSleInc + ClinMotoDec + CommFatig + CommConc
APP =~ NA*ClinAppInc + ClinAppDec + CommAppDec + CommAppInc
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

COGMOODLETH ~~ 1*COGMOODLETH
APP ~~ 1*APP
GATE ~~ 1*GATE
GATE ~~ 0*COGMOODLETH + 0*APP
"
cogmoodleth_app.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=cogmoodleth_app.model, imp_cov=TRUE)

cogmoodleth_app.fit$modelfit
cogmoodleth_app.fit$results[c(1,2,3,6,7, 9)] %>%
   filter(lhs %in% c('COGMOODLETH', 'APP'), rhs %in% c('COGMOODLETH', 'APP'), lhs != rhs) %>%
   mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
   print(digits=2)

```


### Residual correlations

Add residual correlations between same-item symptoms across cohorts. 

```{r cogmoodleth_app_mod, cache=TRUE}

cogmoodleth_app_mod.model <- "
COGMOODLETH =~ NA*CommDep + CommAnh + UkbDep + UkbAnh + ClinSleDec + CommSleDec + ClinMotoInc + ClinGuilt + CommGuilt + ClinSui + CommSui + ClinSleInc + ClinMotoDec + CommSleInc + CommFatig + CommConc
APP =~ NA*CommAppDec + CommAppInc + ClinAppDec + ClinAppInc
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

COGMOODLETH ~~ 1*COGMOODLETH
APP ~~ 1*APP
GATE ~~ 1*GATE
GATE ~~ 0*COGMOODLETH + 0*APP

ClinAppDec ~~ CommAppDec
ClinAppInc ~~ CommAppInc
ClinSleDec ~~ CommSleDec
ClinSleInc ~~ CommSleInc
ClinGuilt ~~ CommGuilt
ClinSui ~~ CommSui
"

cogmoodleth_app_mod.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=cogmoodleth_app_mod.model, imp_cov=TRUE)

cogmoodleth_app_mod.fit$modelfit
cogmoodleth_app_mod.fit$results[c(1,2,3,6,7,9)] %>%
   filter(rhs %in% c('COGMOODLETH', 'APP'), lhs != rhs) %>%
   mutate(STD_Genotype_SE=as.numeric(STD_Genotype_SE)) %>%
   print(digits=2)
```

Check direction of loadings to interpret factor correlations

```{r}
cog_app_veg.fit$results[c(1,2,3,6,7,9)] %>%
  filter(lhs == 'APP')

cogmoodleth_app_mod.fit$results[c(1,2,3,6,7,9)] %>%
  filter(lhs == 'APP')
```

### Model coefficients

Add modified model to list
```{r mod_list}
mod_list <- model_list
mod_list[["M"]] <- list(abbrev = "CogMoodLeth-App", name="Cog/Mood/Leth-Appetite", model=cogmoodleth_app.fit)
mod_list[["N"]] <- list(abbrev = "CogMoodLeth-App [Res]", name="Cog/Mood-Appetite-Lethargy residual correlations", model=cogmoodleth_app_mod.fit)
```

Check that all symptoms have been included in each model

```{r model_check}
sapply(mod_list, function(m) all(names(symptoms_S_var) %in% m$model$results$rhs))
```

Refine naming scheme for publication.
```{r model_coefs}

# mod_list[["Meta"]] <- list(abbrev = "Meta", name="Meta-analysis", model=gwas_meta.fit)

model_abbrevs <- bind_rows(lapply(mod_list, function(m) tibble(abbrev=m$abbrev)), .id='model')
model_names <- bind_rows(lapply(mod_list, function(m) tibble(name=m$name)), .id='model')

model_coefs <- model_abbrevs |>
  left_join(model_names, by = 'model') |>
  left_join(bind_rows(lapply(mod_list, function(m) m$model$results), .id='model'),
            by = 'model') |>
  mutate(lhs = case_when(str_detect(lhs, "Clin") ~ str_replace(lhs, "Clin", "Case"),
                         lhs == "CLIN" ~ "CASE",
                         .default = lhs),
        rhs = case_when(str_detect(rhs, "Clin") ~ str_replace(rhs, "Clin", "Case"),
                        rhs == "CLIN" ~ "CASE",
                        .default = rhs))

write_tsv(model_coefs, "mdd-symptom-gsem-model_files/model_coefs.txt")

```

Get all model fit statistics

```{r}

model_fits_all <- 
tibble(Model=sapply(mod_list, function(m) m$abbrev),
           Name=sapply(mod_list, function(m) m$name)) %>%
bind_cols(
bind_rows(
lapply(mod_list, function(m) m$model$modelfit)
)) %>%
mutate(useAIC = if_else(row_number() <= 10, true = AIC, false = NA)) %>%
mutate(dAIC = useAIC - min(useAIC, na.rm = TRUE)) %>%
select(-useAIC)

write_tsv(model_fits_all, "mdd-symptom-gsem-model_files/model_fits.txt")

```

#### Coefficient plots

Heatmap of how often symptoms load on the same factor in symptoms models

```{r symptom_loadings_heatmap}

symptom_model_connectivity <- model_coefs |>
  filter(! lhs %in% c('MDD', 'CASE', 'COMM', 'GATE'), !model %in% c("M", "N")) |>
  transmute(factor=str_c(model, lhs, sep = "_"), rhs, value = 1) |>
  pivot_wider(names_from = factor, values_from = value, values_fill = 0)

symptom_model_connectivity_matrix <- as.matrix(symptom_model_connectivity[,-1])
rownames(symptom_model_connectivity_matrix) <- symptom_model_connectivity$rhs

heatmap(symptom_model_connectivity_matrix %*% t(symptom_model_connectivity_matrix))

```

Plot structure of symptom-focused models

```{r model_loadings}

model_labels <- model_abbrevs |>
  mutate(model_label = if_else(abbrev == "Meta",
                               true = "Multivariate GWAS:",
                               false = str_glue("{abbrev}:")))
                               
factor_labels <- rev(c("MDD" = "Depression",
"CASE" = "Case-only",
"COMM" = "Community",
"PSYCH" = "Psychological",
"COGMOOD" = "Cognitive/Mood",
"COGMOODLETH" = "Cog/Mood/Leth",
"COG" = "Cognitive",
"MOOD" = "Mood", 
"AFFECT" = "Affective",
"AFC" = "Affect/Cognitive",
"SOMA" = "Somatic",
"APP" = "Appetite/Weight",
"LETH" = "Lethargy",
"NEUROVEG" = "Neurovegetative",
"MEL" = "Melancholic",
"ATY" = "Atypical",
"GATE" = "Gating",
"META" = "F1"))

symptom_model_structure <- model_coefs |>
  filter(op == "=~") |>
  filter(!abbrev %in% c("CogMoodLeth-App [Res]", "Meta")) |>
  mutate(cohort = case_when(str_detect(rhs, "Case") ~ "Case-only",
                            str_detect(rhs, "Comm") ~ "Community",
                            str_detect(rhs, "Ukb") ~ "UKB-T"),
         symptom = str_remove(rhs, "(Case|Comm|Ukb)"),
         loading = if_else(abs(STD_Genotype) > 1, true = 1 * sign(STD_Genotype), false = STD_Genotype),
         Model = factor(model, levels = pull(model_labels, model), labels = pull(model_labels, model_label)),
         Factor = factor(lhs, levels = names(factor_labels), labels = factor_labels))
         
ggplot(symptom_model_structure, aes(x = symptom, y = Factor, colour = cohort, group = cohort, size = abs(loading))) +
  geom_point(position = position_dodge(width = 0.4)) +
  facet_grid(rows = vars(Model), scales = "free", space = "free", switch = "y") +
  scale_x_discrete("Symptom", limits = c("Sui", "Dep", "Anh", "Guilt", "Conc",
                                         "MotoInc", "SleDec", "AppDec", "AppInc",
                                         "MotoDec", "Fatig", "SleInc")) +
  scale_y_discrete('') +
  scale_size("|Loading|", breaks = c(0.1, 0.5, 1)) +
  scale_colour_discrete("Cohorts") +
  theme_minimal() +
  theme(strip.placement = "outside",
        strip.text.y.left = element_text(angle = 0, hjust = 1),
        legend.position = "top",
        legend.spacing.x = unit(1, "pt"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(colour = 'black', linewidth=0.2))

ggsave("mdd-symptom-gsem-model_files/model_loadings.png", width=10, height=8, dpi=300)
ggsave("mdd-symptom-gsem-model_files/model_loadings.pdf", width=10, height=8)
```

#### Model Implied and residual genetic correlations

Plot the implied genetic and residual correlations, but scaled by the total genetic correlation.

```{r}

fit_cor <- function(fit) {
	# get the implied and residual covariances
	cov_imp <-fit$resid_cov$`Model Implied Covariance Matrix`
	cov_res <- fit$resid_cov$`Residual Covariance Matrix: Calculated as Observed Cov - Model Implied Cov`
	
	# get the residual variances from the model
	symp <- fit$results %>% filter(lhs == rhs) %>% pull(lhs)
	res <- fit$results %>% filter(lhs == rhs) %>% pull(Unstand_Est)
	names(res) <- symp
	
	# convert the implied covariance matrix to a correlation matrix
	cor_imp_var <- cov2cor(cov_imp)
	
	# subtract proportion of residual variance from the
	# correlation matrix
	highpass1 <- function(x) if_else(x > 1, true=1, false=x)
	cov_var <- diag(cov_imp)
	diag(cor_imp_var) <- highpass1((cov_var - res[names(cov_var)]) / cov_var)
	
	# do same for residual matrix
	cov_res_var <- cov_res
	diag(cov_res_var) <- diag(cov_imp)
	
	# convert to correlation matrix
	cor_res_var <- cov2cor(cov_res_var)
	
	# replace diagonals
	diag(cor_res_var) <- res[names(cov_var)] / cov_var
	
	return(list(imp_cor=cor_imp_var, res_cor=cor_res_var))
}

```{r mdd_symptoms_d_imp, dpi=300}

d_fit_cor <- fit_cor(measure.fit)

corrplot(d_fit_cor$imp_cor, is.corr=FALSE, col.lim=c(-1, 1))

```

```{r mdd_symptoms_d_resid, dpi=300}

corrplot(d_fit_cor$res_cor, is.corr=FALSE, col.lim=c(-1, 1))

```

```{r mdd_symptoms_i_imp, dpi=300}

i_fit_cor <- fit_cor(cog_app_veg.fit)

corrplot(i_fit_cor$imp_cor, is.corr=FALSE, col.lim=c(-1, 1))

```

```{r mdd_symptoms_i_resid, dpi=300}

corrplot(i_fit_cor$res_cor, is.corr=FALSE, col.lim=c(-1, 1))

```

      
# Exploratory factor analysis

Get the genetic covariance matrix for symptoms with a positive heritability

```{r mdd_symptoms_cov}

symptoms_cov <- symptoms_covstruct$S
k <- nrow(symptoms_cov)
symptoms_se <- matrix(0, k, k)
symptoms_se[lower.tri(symptoms_se, diag=TRUE)] <- sqrt(diag(symptoms_covstruct$V))

symptoms_se[upper.tri(symptoms_se)] <- t(symptoms_se)[upper.tri(symptoms_se)]

symptoms_cov_keep <- which(diag(symptoms_cov > 0))

symptoms_cov_pos <- symptoms_cov[symptoms_cov_keep,symptoms_cov_keep]

```

Smooth the genetic covariance matrix so that it is positive definite
```{r mdd_symptom_gsem_efa_pd, dpi=300}

# smooth the covariance matrix
symptoms_cov_pd <- as.matrix(Matrix::nearPD(symptoms_covstruct$S, corr=FALSE)$mat)

corrplot(cov2cor(symptoms_cov_pd))

```


Check eigen values of the correlation matrix
```{r mdd_symptom_gsem_efa_eigen}

symptoms_eigen <- eigen(cov2cor(symptoms_cov_pd)) 

plot(symptoms_eigen$values, ylab='Eigenvalue')
lines(symptoms_eigen$values)
abline(1, 0, col='red')

symptoms_efa <- factanal(covmat=symptoms_cov_pd, factors=3, rotation='varimax')
print(symptoms_efa, cut=0.4)

```


