---
title: UKB MDD symptom phenotypes 
author: Mark Adams 
output:
  html_document:
    toc: TRUE
    code_folding: hide
    number_sections: TRUE
    df_print: kable
    keep_md: true
  md_document:
    variant: markdown_github
---

# Data extraction

Extract data from the UKB release files

```{bash, eval=FALSE}

cd sumstats/UKB
ukb_unpack ukb12345.enc
wget  -nd  biobank.ndph.ox.ac.uk/showcase/util/encoding.ukb
ukb_conv ukb12345.enc_ukb r -oukb_mdd -iukb_mdd.field 

```

```{r ukb_mdd_r, cache=TRUE}

source('sumstats/UKB/ukb_mdd.r')

```

```{r}

library(dplyr)
library(readr)
library(psych)
library(corrplot)
library(lavaan)

```

# Fields

## DSM lifetime symptoms (CIDI)

Depression symptoms from the Composite International Diagnostic Interview.

```{r}

mhq_cidi_fields <-
read_delim("
MDD1;f.20446.0.0 ;Ever had prolonged feelings of sadness or depression
MDD2;f.20441.0.0 ;Ever had prolonged loss of interest in normal activities
MDD3;f.20536.0.0 ;Weight change during worst episode of depression
MDD4;f.20532.0.0 ;Did your sleep change
MDD4a;f.20533.0.0;Trouble falling asleep
MDD4a;f.20535.0.0;Waking too early
MDD4b;f.20534.0.0;Sleeping too much
MDD6;f.20449.0.0 ;Feelings of tiredness during worst episode of depression
MDD7;f.20450.0.0 ;Feelings of worthlessness during worst period of depression
MDD8;f.20435.0.0 ;Difficulty concentrating during worst depression
MDD9;f.20437.0.0 ;Thoughts of death during worst depression
", col_names=c('Reference', 'Field', 'Description'), delim=';')

mhq_cidi_fields


```


## Current symptoms (PHQ)

Depression symptoms from the Patient Health Questionnaire. 

```{r}

mhq_phq_fields <- 
read_delim("
MDD1;f.20510.0.0;Recent feelings of depression
MDD2;f.20514.0.0;Recent lack of interest or pleasure in doing things
MDD3;f.20511.0.0;Recent poor appetite or overeating
MDD4;f.20517.0.0;Trouble falling or staying asleep or sleeping too much
MDD5;f.20518.0.0;Recent changes in speed amount of moving or speaking
MDD6;f.20519.0.0;Recent feelings of tiredness or low energy
MDD7;f.20507.0.0;Recent feelings of inadequacy
MDD8;f.20508.0.0;Recent trouble concentrating on things
MDD9;f.20513.0.0;Recent thoughts of suicide or self harm
", col_names=c('Reference', 'Field', 'Description'), delim=';')

mhq_phq_fields

```
   
   
   
# Input coding

## CIDI
   

Code up different kinds of missingness (did not take MHQ vs declind screening symptoms vs answered negative to screening symptoms  vs declined symptom)

```{r}

# Symptoms coded with codings 502 and 503
# Coding 503 for Criterion A1/A2
## -818	Prefer not to answer
##    0	No
##    1	Yes
#
# 'Not assessed' = not assessed by MHQ
# 'Declined'     = preferred not to answer
# 'Absent' = reports not having symptom 
# 'Present'  = reports having symptom 

# Coding 502 for other Criteria (except weight change)
## -818	Prefer not to answer
## -121	Do not know
##    0	No
##    1	Yes

# 'Not assessed' = not assessed by MHQ
# 'Missing'      = did not respond to depression/anhedonia items
# 'Screened'     = answered 0 for depression and anhedonia
# 'Declined'     = did not know or preferred not to answer
# 'Absent' = reports not having symptom 
# 'Present'  = reports having symptom 

resp_prefer <- 'Prefer not to answer'
resp_know <- 'Do not know'

screen_symptom_502 <- function(symptom_field, depression_field, anhedonia_field) {
        case_when(symptom_field %in% c(resp_prefer, resp_know) ~ 'Declined',
                  symptom_field %in% 'No' ~ 'Absent',
                  symptom_field %in% 'Yes' ~ 'Present',
                  depression_field %in% 'No' | anhedonia_field %in% 'No' ~ 'Screened',
                  depression_field %in% resp_prefer & anhedonia_field %in% resp_prefer ~ 'Missing',
                  TRUE ~ 'Not assessed')
}

# Coding for sleep criteria

### f.20534.0.0		Sleeping.too.much									###
### f.20533.0.0		Trouble.falling.asleep									###
### f.20535.0.0		Waking.too.early									###
# Coding 508
## 0	No
## 1	Yes

# Coding 507 for weight change criterion
# 
## -818	Prefer not to answer
## -121	Do not know
##    0	Stayed about the same or was on a diet
##    1	Gained weight
##    2	Lost weight
##    3	Both gained and lost some weight during the episode

# 'Not assessed' = not assessed by MHQ
# 'Missing'      = did not respond to depression/anhedonia items
# 'Screened'     = answered 0 for depression and anhedonia
# 'Declined'     = did not know or preferred not to answer
# 'Absent'       = reports not having symptom 
# 'Present'      = reports having symptom 

resp_same <- 'Stayed about the same or was on a diet'
resp_gain <- 'Gained weight'
resp_lost <- 'Lost weight'
resp_both <- 'Both gained and lost some weight during the episode'


screen_symptom_507 <- function(symptom_field, depression_field, anhedonia_field) {
        case_when(symptom_field %in% c(resp_prefer, resp_know) ~ 'Declined',
                  symptom_field %in% 'No' ~ 'Absent',
                  symptom_field %in% c(resp_gain, resp_lost, resp_both) ~ 'Present',
                  depression_field %in% 'No' | anhedonia_field %in% 'No' ~ 'Screened',
                  depression_field %in% resp_prefer & anhedonia_field %in% resp_prefer ~ 'Missing',
                  TRUE ~ 'Not assessed')
}

# code weight 'gain'/'loss' as a symptom
screen_weight_507 <- function(symptom_field, present_levels, absent_levels, depression_field, anhedonia_field) {
        case_when(symptom_field %in% c(resp_prefer, resp_know) ~ 'Declined',
                  symptom_field %in% absent_levels ~ 'Absent',
                  symptom_field %in% present_levels ~ 'Present',
                  depression_field %in% 'No' | anhedonia_field %in% 'No' ~ 'Screened',
                  depression_field %in% resp_prefer & anhedonia_field %in% resp_prefer ~ 'Missing',
                  TRUE ~ 'Not assessed')
}

```  

```{r}

CIDI_symptoms <- bd %>%
select(f.eid, f.20446.0.0, f.20441.0.0, f.20532.0.0, f.20534.0.0, f.20533.0.0, f.20517.0.0, f.20535.0.0, f.20435.0.0, f.20449.0.0, f.20450.0.0, f.20437.0.0, f.20536.0.0) %>%
mutate(
# Depression and anhedonia act as screening items for the rest of the symptoms
       cidi1=case_when(is.na(f.20446.0.0) ~ 'Not assessed',
                                 f.20446.0.0 == resp_prefer ~ 'Missing',
                                 f.20446.0.0 == 'No' ~ 'Absent',
                                 f.20446.0.0 == 'Yes' ~ 'Present'),
       cidi2=case_when(is.na(f.20441.0.0) ~ 'Not assessed',
                                 f.20441.0.0 == resp_prefer ~ 'Missing',
                                 f.20441.0.0 == 'No' ~ 'Absent',
                                 f.20441.0.0 == 'Yes' ~ 'Present'),
# Weight change
       cidi3=screen_weight_507(symptom_field=f.20536.0.0,
                               present_levels=c(resp_gain, resp_lost, resp_both),
                               absent_levels=c(resp_same),
                               depression_field=f.20446.0.0, anhedonia_field=f.20441.0.0),
# Weight gain
       cidi3b=screen_weight_507(symptom_field=f.20536.0.0,
                                    present_levels=c(resp_gain, resp_both),
                                    absent_levels=c(resp_same, resp_lost),
                                    depression_field=f.20446.0.0, anhedonia_field=f.20441.0.0),
# Weight loss
       cidi3a=screen_weight_507(symptom_field=f.20536.0.0,
                                    present_levels=c(resp_lost, resp_both),
                                    absent_levels=c(resp_same, resp_gain),
                                    depression_field=f.20446.0.0, anhedonia_field=f.20441.0.0),
# Sleep (insomnia or hypersomnia)
       cidi4=screen_symptom_502(symptom_field=f.20532.0.0,
                                     depression_field=f.20446.0.0, anhedonia_field=f.20441.0.0),
### f.20534.0.0		Sleeping.too.much					 
### f.20533.0.0		Trouble.falling.asleep		 
### f.20535.0.0		Waking.too.early					 
# Coding 508
## 0	No
## 1	Yes
# Insomnia
       cidi4a=case_when(f.20532.0.0 %in% c(resp_prefer, resp_know) ~ 'Declined',
                               f.20533.0.0 %in% 'Yes' | f.20535.0.0 %in% 'Yes' ~ 'Present',
                               f.20534.0.0 %in% 'Yes' ~ 'Absent',
                               f.20532.0.0 %in% 'No' ~ 'Absent',
                               f.20446.0.0 %in% 'No' | f.20441.0.0 %in% 'No' ~ 'Screened',
                               f.20446.0.0 %in% resp_prefer & f.20441.0.0 %in% resp_prefer ~ 'Missing',
                               TRUE ~ 'Not assessed'),
# Hypersomnia
       cidi4b=case_when(f.20532.0.0 %in% c(resp_prefer, resp_know) ~ 'Declined',
                                  f.20534.0.0 %in% 'Yes' ~ 'Present',
                                  f.20533.0.0 %in% 'Yes' | f.20535.0.0 %in% 'Yes' ~ 'Absent',
                                  f.20532.0.0 %in% 'No' ~ 'Absent',
                                  f.20446.0.0 %in% 'No' | f.20441.0.0 %in% 'No' ~ 'Screened',
                                  f.20446.0.0 %in% resp_prefer & f.20441.0.0 %in% resp_prefer ~ 'Missing',
                                  TRUE ~ 'Not assessed'),
# Tiredness
       cidi6=screen_symptom_502(symptom_field=f.20449.0.0,
                                     depression_field=f.20446.0.0, anhedonia_field=f.20441.0.0),
# Worthlessness
       cidi7=screen_symptom_502(symptom_field=f.20450.0.0,
                                     depression_field=f.20446.0.0, anhedonia_field=f.20441.0.0),
# Concentration
       cidi8=screen_symptom_502(symptom_field=f.20435.0.0,
                                     depression_field=f.20446.0.0, anhedonia_field=f.20441.0.0),
# Thoughts of death
       cidi9=screen_symptom_502(symptom_field=f.20437.0.0,
                                     depression_field=f.20446.0.0, anhedonia_field=f.20441.0.0))


``` 

Code to binary (absent = 0, present = 1)

```{r}

Ab0Pr1 <- function(symptom) case_when(symptom == 'Absent' ~ 0L,
                                    symptom == 'Present' ~ 1L,
                                    TRUE ~ NA_integer_)

CIDI_binary <- CIDI_symptoms %>%
mutate(CIDI1=Ab0Pr1(cidi1),
       CIDI2=Ab0Pr1(cidi2),
       CIDI3=Ab0Pr1(cidi3),
       CIDI3a=Ab0Pr1(cidi3a),
       CIDI3b=Ab0Pr1(cidi3b),
       CIDI4=Ab0Pr1(cidi4),
       CIDI4a=Ab0Pr1(cidi4a),
       CIDI4b=Ab0Pr1(cidi4b),
       CIDI6=Ab0Pr1(cidi6),
       CIDI7=Ab0Pr1(cidi7),
       CIDI8=Ab0Pr1(cidi8),
       CIDI9=Ab0Pr1(cidi9))

```

## PHQ

Question form: "Over the last 2 weeks, how often have you been bothered by any of the following problems?"

- "Feeling down, depressed, or hopeless"
- etc.

```{r}

# Coding
# -818	Prefer not to answer
# 1	Not at all
# 2	Several days
# 3	More than half the days
# 4	Nearly every day

# recode responses as integers 0:3

recode_phq <- function(x) recode(x, `Not at all`=0L,
                                    `Several days`=1L,
                                    `More than half the days`=2L,
                                    `Nearly every day`=3L,
                                    .default=NA_integer_)

phq <- bd %>%
select(f.eid, f.20510.0.0, f.20514.0.0, f.20511.0.0, f.20517.0.0, f.20518.0.0, f.20519.0.0, f.20507.0.0, f.20508.0.0, f.20513.0.0) %>%
mutate(PHQ1=recode_phq(f.20510.0.0),
       PHQ2=recode_phq(f.20514.0.0),
       PHQ3=recode_phq(f.20511.0.0),
       PHQ4=recode_phq(f.20517.0.0),
       PHQ5=recode_phq(f.20518.0.0),
       PHQ6=recode_phq(f.20519.0.0),
       PHQ7=recode_phq(f.20507.0.0),
       PHQ8=recode_phq(f.20508.0.0),
       PHQ9=recode_phq(f.20513.0.0))

```

# Phenotypic correlations

Calculate tetrachoric correlations for binary data

```{r ukb_pheno_cidi_phq_cor}

cidi_phq <- CIDI_binary %>%
inner_join(phq, by='f.eid') %>%
select(starts_with('CIDI', ignore.case=FALSE),
       starts_with('PHQ', ignore.case=FALSE))


options('mc.cores'=4)

cidi_phq_cor <- mixedCor(cidi_phq, p=13:21, d=1:12)

corrplot(cidi_phq_cor$rho, 'square')

dput(cidi_phq_cor, 'sumstats/UKB/ukb_cidi_phq_mixed_cor.deparse.R', control=c('exact'))


```

# Confirmatory models

```{r}

cidi.model <- "F1 =~ NA*CIDI1 + CIDI2 + CIDI3a + CIDI3b + CIDI4a + CIDI4b + CIDI6 + CIDI7 + CIDI8 + CIDI9
F1 ~~ 1*F1
"

cidi.fit <- cfa(cidi.model, data=cidi_phq)

summary(cidi.fit)

fitMeasures(cidi.fit)

```
