---
title: GenomicSEM of MDD symptom factors with other phenotypes setup
author: Mark Adams, Bradley Jermy, Jackson Thorp, Andrew Grotzinger, Michel Nivard 
output:
  html_document:
    toc: TRUE
    code_folding: hide
    number_sections: TRUE
    df_print: kable
    keep_md: true
  md_document:
    variant: markdown_github
---

Test genetic relationship between [symptom factors](mdd-symptom-gsem-model.md) and a selection of other phenotypes that are genetically correlated with MDD, as well as to MDD itself. Phenotypes to examine:

- Major depressive disorder: Clinical cohorts from [Wray et al](https://www.nature.com/articles/s41588-018-0090-3%5C) and all cohorts from [Howard et al](https://www.nature.com/articles/s41593-018-%200326-7). Download from [PGC](https://www.med.unc.edu/pgc/download-results/) and obtain via [data access](https://www.med.unc.edu/pgc/shared-methods/how-to/).
- bipolar disorder: [Mullins et al](https://pubmed.ncbi.nlm.nih.gov/34002096/). Download from [PGC](https://figshare.com/articles/dataset/PGC3_bipolar_disorder_GWAS_summary_statistics/14102594)
- anxiety: [meta-analysis](https://drive.google.com/drive/folders/1fguHvz7l2G45sbMI9h_veQun4aXNTy1v) of [UKBB, iPSYCH](https://www.nature.com/articles/s41380-019-0559-1), and [ANGST](https://pubmed.ncbi.nlm.nih.gov/26754954/), from [Grotzinger et al medRxiv](https://www.medrxiv.org/content/10.1101/2020.09.22.20196089v1.full)
- PTSD: [Nievergelt et al](https://pubmed.ncbi.nlm.nih.gov/31594949/). Download from [PGC](https://figshare.com/articles/dataset/ptsd2019/14672133)
- tobacco use. Cigarettes per day [Liu et al](https://www.nature.com/articles/s41588-018-0307-5). Download from [UofM](https://conservancy.umn.edu/handle/11299/201564)
- alcohol dependence. [Walters et al](https://www.nature.com/articles/s41593-018-0275-1). Download from [PGC](https://doi.org/10.6084/m9.figshare.14672187)
- educational attainment. [Okbay et. al](https://www.nature.com/articles/s41588-022-01016-z). Download from [SSGAC Data Portal](https://thessgac.com).
- BMI, sex combined [Pulit et al](https://academic.oup.com/hmg/article/28/1/166/5098227). Download from [GIANT/Broad](https://portals.broadinstitute.org/collaboration/giant/index.php/GIANT_consortium_data_files#2018_GIANT_and_UK_BioBank_Meta-analysis).
- neuroticism: [Nagel et al](https://www.nature.com/articles/s41588-018-0151-7). Download from [CNCR](https://ctg.cncr.nl/software/summary_statistics)
- pain: multisite chronic pain [Johnston et al](https://journals.plos.org/plosgenetics/article?id=10.1371/journal.pgen.1008164). Download from [UoG](https://researchdata.gla.ac.uk/822/)
- chronotype: long sleep duration. [Dashti et al](https://www.ncbi.nlm.nih.gov/pubmed/30846698). Download from [SDKP](https://sleep.hugeamp.org/downloads.html).

# Setup

## R packages

R version

```{r version}

R.version

```

Package installation

```{r install, output='hide', warning=FALSE, message=FALSE, eval=FALSE}

required_packages <- c('devtools', 'readr', 'tidyr', 'dplyr', 'ggplot2', 'stringr', 'corrplot')
for(pack in required_packages) if(!require(pack, character.only=TRUE)) install.packages(pack)

library(devtools)

if(!require(GenomicSEM)) install_github("MichelNivard/GenomicSEM")

if(!require(tidySEM)) install_github("cjvanlissa/tidySEM")

```

GenomicSEM version
```{r require, warning=FALSE, messages=FALSE}

require(readr)
require(tidyr)
require(stringr)
require(dplyr)
require(ggplot2)
require(corrplot)
require(GenomicSEM)

packageVersion("GenomicSEM")

```

# Process external sumstats

## Reformat

Format the external sumstats for reading by GenomicSEM with columns `SNP`, `A1` (effect allele), `A2` (non-effect allele), `BETA`/`OR`, `P`, `INFO`, and `N`.

```{r ext2txt, eval=FALSE}

# Major depressive disorder
mdd <- read_table('sumstats/PGC_UKB_23andMe_depression_genome-wide_info_N.txt.gz')
mdd_sumstats <- mdd %>%
    mutate(Nca=UKB_Ncases+PGC_Ncases+X23andMe_Ncases,
           Nco=UKB_Ncontrols+PGC_Ncontrols+X23andMe_Ncontrols) %>%
    transmute(SNP=MarkerName, A1=toupper(Allele1), A2=toupper(Allele2),
              BETA=Effect, SE=StdErr, P=P.value, 
              FREQ=Freq1, N=4*Nca*Nco/(Nca+Nco))
write_tsv(mdd_sumstats, 'sumstats/MD.txt')

mdd_clin <- read_tsv('sumstats/daner_MDD29.0515a_mds6.0316.gz')
mdd_clin_sumstats <- mdd_clin %>%
    transmute(SNP, A1, A2, BETA=log(OR), SE, FREQ=FRQ_U_25632, INFO, P,
              N=4*Nca*Nco/(Nca+Nco))
write_tsv(mdd_clin_sumstats, 'sumstats/MDD.txt')

# Bipolar disorder
bip <- read_tsv('sumstats/pgc-bip2021-all.vcf.tsv.gz', comment='##')
bip_sumstats <- bip %>%
    filter(IMPINFO >= 0.6) %>%
    transmute(SNP=ID, A1, A2, BETA, SE, P=PVAL, INFO=IMPINFO, N=2*NEFFDIV2)
write_tsv(bip_sumstats, 'sumstats/BIP.txt')

# alcohol dependence
alcdep <- read_table('sumstats/pgc_alcdep.eur_discovery.aug2018_release.txt.gz')
alcdep_sumstats <- alcdep %>%
mutate(SNP=str_split_fixed(SNP, pattern=":", n=2)[,1]) %>%
select(SNP, A1, A2, BETA=Z, P, N=Weight)
write_tsv(alcdep_sumstats, 'sumstats/AlcDep.txt')

# body-mass index
bmi <- read_table('sumstats/Bmi.giant-ukbb.meta-analysis.combined.23May2018.HapMap2_only.txt.gz')
bmi_sumstats <- bmi %>%
    separate(SNP, into=c('SNP', 'A1', 'A2'), sep=':') %>%
    filter(INFO >= 0.6) %>%
    select(SNP, A1=Tested_Allele, A2=Other_Allele, BETA, SE, P, INFO, N)
write_tsv(bmi_sumstats, 'sumstats/BMI.txt')

# educational attainment
ea <- read_tsv('sumstats/EA4_additive_excl_23andMe.txt.gz')
ea_sumstats <-  ea %>%
    transmute(SNP=rsID, A1=Effect_allele, A2=Other_allele, FREQ=EAF_HRC, BETA=Beta, SE_unadj, P=P_unadj, N=765283)
write_tsv(ea_sumstats, 'sumstats/EA.txt')

# multisite chronic pain
pain <- read_tsv('sumstats/chronic_pain-bgen.stats.gz')
pain_sumstats <- pain %>%
    filter(INFO >= 0.6) %>%
    transmute(SNP, A1=ALLELE1, A2=ALLELE0, BETA, SE, P=P_BOLT_LMM_INF, INFO, N=387649) 
write_tsv(pain_sumstats, 'sumstats/Pain.txt')

# tabacco use
smoking <- read_table('sumstats/CigarettesPerDay.txt.gz')
smoking_sumstats <- smoking %>%
    select(SNP=RSID, A1=ALT, A2=REF, BETA, SE, P=PVALUE, N)
write_tsv(smoking_sumstats, 'sumstats/Smoking.txt')

# chronotype (long sleep duration)
sleep <- read_table('sumstats/longsumstats.txt.gz')
sleep_sumstats <- sleep %>%
    filter(INFO >= 0.6) %>%
    transmute(SNP, A1=ALLELE1, A2=ALLELE0, BETA=BETA_LONGSLEEP, SE=SE_LONGSLEEP, P=P_LONGSLEEP, INFO, N=4*34184*(305742-34184)/305742)
write_tsv(sleep_sumstats, 'sumstats/Sleep.txt')

# anxiety disorder
anxiety <- read_table('sumstats/META_UKBB_iPSYCH_ANGST_wNcol.sumstats.gz')
anxiety_sumstats <- anxiety %>%
    mutate(Nca=25453+12655+7016, Nco=58113+19225+14745) %>%
    transmute(SNP, A1=Allele1, A2=Allele2, BETA=Effect, SE=StdErr, P, N=4*Nca*Nco/(Nca+Nco))
write_tsv(anxiety_sumstats, 'sumstats/Anxiety.txt')    

# post-traumatic stress disorder
ptsd <- read_tsv('sumstats/pts_eur_freeze2_overall.results.gz')
ptsd_sumstats <- ptsd %>%
    transmute(SNP,  A1, A2, OR, SE, P, INFO, N=4*Nca*Nco/(Nca+Nco))
write_tsv(ptsd_sumstats, 'sumstats/PTSD.txt')
    
# neuroticism
neu <- read_tsv('sumstats/sumstats_neuroticism_ctg_format.txt.gz')
neu_sumstats <- neu %>%
    filter(!is.na(INFO_UKB)) %>%
    select(SNP=RSID, A1, A2, BETA=Z, P, INFO=INFO_UKB, N)
write_tsv(neu_sumstats, 'sumstats/Neu.txt')

```

## Munge

```{r ext, eval=TRUE}

ext_traits <- c('AlcDep'=0.159, 'Anxiety'=0.16, 'BIP'=0.01,
                'BMI'=NA, 'EA'=NA, 'MD'=0.3, 'MDD'=0.15, 'Neu'=NA,
                'PTSD'=0.3, 'Pain'=NA, 'Sleep'=0.11, 'Smoking'=NA)
ext_trait_names <- names(ext_traits)

```

```{r extmunge, eval=FALSE}

munge(file.path('sumstats', paste(ext_trait_names, 'txt', sep='.')),
trait.names=ext_trait_names,
hm3=here::here("sumstats/reference/w_hm3.snplist"),
info.filter = 0.9, maf.filter = 0.01)


```

# Symptom labels

MDD DSM symptoms are numbered 1-9:

```{r labels}

# plot labels

dsm_mdd_symptoms_labels <-
read_delim("
MDD1;Mood;Mood;Dep
MDD2;Interest;Interest;Anh
MDD3;Weight⇅;Weight⇆;App
MDD3a;Weight⇊;Weight⇇;AppDec
MDD3b;Weight⇈;Weight⇉;AppInc
MDD4;Sleep⇅;Sleep⇆;Sle
MDD4a;Sleep⇊;Sleep⇇;SleDec
MDD4b;Sleep⇈;Sleep⇉;SleInc
MDD5;Motor⇅;Motor⇆;Moto
MDD5a;Motor⇈;Motor⇉;MotoInc
MDD5b;Motor⇊;Motor⇇;MotoDec
MDD6;Fatigue;Fatigue;Fatig
MDD7;Guilt;Guilt;Guilt
MDD8;Concentrate;Concentrate;Conc
MDD9;Suicidality;Suicidality;Sui
", col_names=c('ref', 'h', 'v', 'abbv'), delim=';')

dsm_mdd_symptoms_reference <-
read_delim("
MDD1;Depressed mood most of the day, nearly every day
MDD2;Markedly diminished interest or pleasure in all, or almost all, activities most of the day, nearly every day
MDD3;Significant change in weight or appetite
MDD3a;Significant weight loss or decrease in appetite
MDD3b;Significant weight gain or increase in appetite
MDD4;Sleeping too much or not sleeping enough
MDD4a;Insomnia nearly every day
MDD4b;Hypersomnia nearly every day
MDD5;Changes in speed/amount of moving or speaking
MDD5a;Psychomotor agitation nearly every day
MDD5b;Psychomotor slowing nearly every day
MDD6;Fatigue or loss of energy nearly every day
MDD7;Feelings of worthlessness or excessive or inappropriate guilt
MDD8;Diminished ability to think or concentrate, or indecisiveness
MDD9;Recurrent thoughts of death or suicide or a suicide attempt or a Multiple plan for attempting suicide
", col_names=c('Reference', 'Description'), delim=';')

```


# Symptom prevalences

Load previously calculated symptom prevalences:

```{r prevs}

all_covstruct_prefix <- 'clin.comm.covstruct'
all_sumstats_prevs <- read_tsv(here::here('ldsc', paste(all_covstruct_prefix, 'prevs', 'txt', sep='.'))) 

```


# Multivariable LDSC estimation

Calculate LDSC covariance structure for syptoms used in the combined structural model and the external phenotypes.

```{r ldsc, results='hide'}

covstruct_prefix <- 'symptoms.external.covstruct'
covstruct_r <- here::here('ldsc', paste(covstruct_prefix, 'deparse.R', sep='.'))
covstruct_rds <- here::here('ldsc', paste(covstruct_prefix, 'rds', sep='.'))

if(!file.exists(covstruct_r)) {
    
  symptoms_sumstats_prevs <- all_sumstats_prevs %>%
  left_join(dsm_mdd_symptoms_labels, by=c('symptom'='ref')) %>%
  mutate(samp_prev=0.5,
         cohort=case_when(str_detect(filename, 'Clin') ~ 'Clin',
                          str_detect(filename, 'Comm') ~ 'Comm',
                          str_detect(filename, 'UKBt') ~ 'Ukb')) %>%
  transmute(filename=here::here(filename), samp_prev, pop_prev, trait_name=paste0(cohort, abbv))

  # external files, prevalences and trait names
  external_sumstats_prevs <-
    tibble(filename=paste(ext_trait_names, 'sumstats', 'gz', sep='.'),
           trait_name=ext_trait_names,
           pop_prev=ext_traits) %>%
    mutate(samp_prev=if_else(!is.na(pop_prev), true=0.5, false=NA_real_))
    
  sumstats_prevs <- bind_rows(symptoms_sumstats_prevs, external_sumstats_prevs)
    

  symptoms_covstruct <- ldsc(traits=sumstats_prevs$filename,
                             sample.prev=sumstats_prevs$samp_prev,
                             population.prev=sumstats_prevs$pop_prev,
                             ld=here::here('sumstats/reference/eur_w_ld_chr/'),
                             wld=here::here('sumstats/reference/eur_w_ld_chr/'),
                             trait.names=sumstats_prevs$trait_name)

  dput(symptoms_covstruct, covstruct_r, control=c('exact'))
  saveRDS(symptoms_covstruct, covstruct_rds)
  
  # check for exact match of deparsed object
  identical(dget(covstruct_r), symptoms_covstruct)

} else {

  symptoms_covstruct <- dget(covstruct_r)

}


```

# Models

## Symptom factors

Base model of symptom factors, with merged Cognitive/Vegeative and Appetite.

```{r symptom_model, cache=TRUE}

model <- "
DEP =~ NA*CommDep + CommAnh + UkbDep + UkbAnh + ClinSleDec + CommSleDec + ClinMotoInc + ClinGuilt + CommGuilt + ClinSui + CommSui + ClinSleInc + CommSleInc + ClinMotoDec + CommFatig + CommConc
APP =~ NA*ClinAppInc + ClinAppDec + CommAppDec + app_co3b*CommAppInc
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

DEP ~~ 1*DEP
APP ~~ 1*APP
GATE ~~ 1*GATE
GATE ~~ 0*DEP + 0*APP
"
fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=model, CFIcalc=TRUE)

```

## Sample factors

Base model of Clinical and Community sample factors. 

```{r sample_model, cache=TRUE}

model <- "
CLIN =~ NA*ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec + ClinGuilt + ClinSui
COMM =~ NA*CommDep + CommAnh + UkbDep + UkbAnh + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig + CommGuilt + CommConc + CommSui
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

CLIN ~~ 1*CLIN
GATE ~~ 1*GATE
COMM ~~ 1*COMM
GATE ~~ 0*COMM + 0*CLIN
"
fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=model, CFIcalc=TRUE)

```

## External phenotypes

### Symptom factors

Compare symptom factors against each external phenotype. Single regression of each external phenotype on each symptom or symptom cluster.

```{r ext_glue, cache=TRUE}

ext_symp.glue <- "
DEP =~ NA*CommDep + CommAnh + UkbDep + UkbAnh + ClinSleDec + CommSleDec + ClinMotoInc + ClinGuilt + CommGuilt + ClinSui + CommSui + ClinSleInc + CommSleInc + ClinMotoDec + CommFatig + CommConc
APP =~ NA*ClinAppInc + ClinAppDec + CommAppDec + app_co3b*CommAppInc
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

DEP ~~ 1*DEP
APP ~~ 1*APP
GATE ~~ 1*GATE
GATE ~~ 0*DEP + 0*APP

AlcDep ~ {symptom}
Anxiety ~  {symptom}
BIP ~ {symptom}
BMI ~ {symptom}
EA ~ {symptom}
MD ~ {symptom}
MDD ~ {symptom}
Neu ~ {symptom}
PTSD ~ {symptom}
Pain ~ {symptom}
Sleep ~ {symptom}
Smoking ~ {symptom}
"

ext_symp.model_list <- lapply(c('DEP', 'APP', 'GATE'), function(symptom) str_glue_data(list(symptom=symptom), ext_symp.glue))

ext_symp.fit_list <- lapply(ext_symp.model_list, function(model) usermodel(symptoms_covstruct, estimation='DWLS', model=model))

```

### Sample factors

```{r ext_samp_glue, cache=TRUE}

ext_samp.glue <- "
CLIN =~ NA*ClinGuilt + ClinSui + ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec
COMM =~ NA*CommDep + CommAnh + UkbDep + UkbAnh + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig + CommGuilt + CommConc + CommSui
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

CLIN ~~ 1*CLIN
GATE ~~ 1*GATE
COMM ~~ 1*COMM
GATE ~~ 0*COMM + 0*CLIN

cc < 0.99
CLIN ~~ cc*COMM

AlcDep ~ {symptom}
Anxiety ~  {symptom}
BIP ~ {symptom}
BMI ~ {symptom}
EA ~ {symptom}
MD ~ {symptom}
MDD ~ {symptom}
Neu ~ {symptom}
PTSD ~ {symptom}
Pain ~ {symptom}
Sleep ~ {symptom}
Smoking ~ {symptom}
"

ext_samp.model_list <- lapply(c('CLIN', 'COMM'), function(symptom) str_glue_data(list(symptom=symptom), ext_samp.glue))

ext_samp.fit_list <- lapply(ext_samp.model_list, function(model) usermodel(symptoms_covstruct, estimation='DWLS', model=model))

```

```{r ext_loadings}

ext_single <-
bind_rows(lapply(c(ext_symp.fit_list, ext_samp.fit_list), function(fit) mutate(fit$results, p_value=as.character(p_value)))) %>%
select(lhs, op, rhs, STD_Genotype, STD_Genotype_SE, p_value) %>%
filter(lhs %in% ext_trait_names, rhs %in% c('DEP', 'APP', 'GATE', 'CLIN', 'COMM')) %>%
mutate(Beta='Single', Factor=rhs, Phenotype=lhs, p_value=as.numeric(p_value))

```

### Symptom factors

Multiple regression of each phenotype on symptom factors, to estimate relationship after condition on each of the other factors. 


```{r ext_mult, cache=TRUE}

ext_mult_symp.model <- "
DEP =~ NA*CommDep + CommAnh + UkbDep + UkbAnh + ClinSleDec + CommSleDec + ClinMotoInc + ClinGuilt + CommGuilt + ClinSui + CommSui + ClinSleInc + CommSleInc + ClinMotoDec + CommFatig + CommConc
APP =~ NA*ClinAppInc + ClinAppDec + CommAppDec + CommAppInc
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

DEP ~~ 1*DEP
APP ~~ 1*APP
GATE ~~ 1*GATE
GATE ~~ 0*DEP + 0*APP

AlcDep ~ DEP + APP + GATE
Anxiety ~  DEP + APP + GATE
BIP ~ DEP + APP + GATE
BMI ~ DEP + APP + GATE
EA ~ DEP + APP + GATE
MD ~ DEP + APP + GATE
MDD ~ DEP + APP + GATE
Neu ~ DEP + APP + GATE
PTSD ~ DEP + APP + GATE
Pain ~ DEP + APP + GATE
Sleep ~ DEP + APP + GATE
Smoking ~ DEP + APP + GATE
"
ext_mult_symp.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=ext_mult_symp.model)

```

### Sample factors


```{r ext_samp_mult, cache=TRUE}

ext_samp_mult.model <- "
CLIN =~ NA*ClinGuilt + ClinSui + ClinAppDec + ClinAppInc + ClinSleDec + ClinSleInc + ClinMotoInc + ClinMotoDec
COMM =~ NA*CommDep + CommAnh + UkbDep + UkbAnh + CommAppDec + CommAppInc + CommSleDec + CommSleInc + CommFatig + CommGuilt + CommConc + CommSui
GATE =~ NA*CommDep + CommAnh + UkbDep + UkbAnh

CLIN ~~ 1*CLIN
GATE ~~ 1*GATE
COMM ~~ 1*COMM
GATE ~~ 0*COMM + 0*CLIN

AlcDep ~ CLIN + COMM 
Anxiety ~  CLIN + COMM
BIP ~ CLIN + COMM 
BMI ~ CLIN + COMM 
EA ~ CLIN + COMM 
MD ~ CLIN + COMM
MDD ~ CLIN + COMM
Neu ~ CLIN + COMM
PTSD ~ CLIN + COMM
Pain ~ CLIN + COMM
Sleep ~ CLIN + COMM
Smoking ~ CLIN + COMM
"
ext_samp_mult.fit <- usermodel(symptoms_covstruct, estimation='DWLS', model=ext_samp_mult.model)

```

```{r ext_mult_loadings}

ext_multiple <-
bind_rows(lapply(list(ext_mult_symp.fit, ext_samp_mult.fit),
                 function(fit) fit$results)) %>%
  select(lhs, op, rhs, STD_Genotype, STD_Genotype_SE, p_value) %>%
  filter(lhs %in% ext_trait_names, rhs %in% c('DEP', 'APP', 'GATE', 'CLIN', 'COMM')) %>%
  mutate(Beta='Multiple', Factor=rhs, Phenotype=lhs)

```

```{r ext_table}

ext_table <-
bind_rows(ext_single, ext_multiple) %>%
transmute(Phenotype,
          Factor=factor(Factor,
	    				levels=c('DEP', 'APP', 'GATE', 'CLIN', 'COMM'),
						labels=c("Depression", "Appetite/Weight", "Gating", "Community", "Clinical")),
		  Model=Beta, STD_Genotype, STD_Genotype_SE, p_value) %>%
mutate(fdr=p.adjust(p_value, 'BY')) |>
arrange(Phenotype, desc(Factor), Model)

ext_table

```

```{r ex_plot, fig.width=10, fig.height=6, dpi=300}

ggplot(bind_rows(ext_multiple, ext_single),
       aes(x=factor(Factor,
	   				levels=c('DEP', 'APP', 'GATE', 'CLIN', 'COMM'),
					   labels=c("Depression", "Appetite/Weight", "Gating", "Community", "Clinical")),
           y=STD_Genotype,
           color=factor(Beta, levels=c('Single', 'Multiple')),
           shape=factor(Beta, levels=c('Single', 'Multiple')),
          ymin=qnorm(0.025, mean=STD_Genotype, sd=as.numeric(STD_Genotype_SE)),
          ymax=qnorm(0.975, mean=STD_Genotype, sd=as.numeric(STD_Genotype_SE)))) +
geom_hline(yintercept=0, col='gray') +
geom_pointrange(position=position_dodge(width=-0.5)) +
facet_wrap(~Phenotype) +
scale_x_discrete('Factor') +
scale_y_continuous('Beta') +
scale_color_discrete('Regression: ') +
coord_flip(ylim=c(-1.5, 1.5)) +
theme_bw() +
theme(axis.text.y=element_text(size=16),
      strip.text=element_text(size=16),
      legend.title=element_text(size=12),
      legend.text=element_text(size=14),
      legend.position='top') +
labs(color  = "Regression: ", shape = "Regression: ")

```

Consistancy of effect direction in single regression

```{r ext_direction}
ext_table |>
filter(Model == 'Single') |>
group_by(Phenotype) |>
filter(any(sign(STD_Genotype) == 1) & any(sign(STD_Genotype) == -1))
```

Specific relationship
```{r specific_sample}
specific_sample <-
ext_table |>
filter(Factor %in% c('Clinical', 'Community'), Model == 'Multiple') |>
group_by(Phenotype) |>
filter(any(fdr <= 0.05), any(fdr > 0.05))  |>
filter(fdr <= 0.05)
specific_sample
```

```{r specific_symptoms}
ext_table |>
filter(Factor %in% c('Appetite/Weight', 'Depression', 'Gating'), Model == 'Multiple') |>
group_by(Phenotype) |>
filter(any(fdr <= 0.05), any(fdr > 0.05)) |>
ungroup() |>
filter(fdr <= 0.05) |>
arrange(Factor)
```

Shared relationships
```{r shared_sample}
ext_table |>
filter(Factor %in% c('Clinical', 'Community')) |>
group_by(Phenotype) |>
filter(any(fdr < 0.05)) |>
filter(Model == 'Multiple') |>
filter(all(fdr >= 0.05))
```
```{r shared_symptoms}
ext_table |>
filter(Factor %in% c('Appetite/Weight', 'Depression', 'Gating'), Model == 'Multiple') |>
group_by(Phenotype) |>
filter(all(fdr <= 0.05))
```

Test for attenuation of correlations

```{r ext_attentuate}

ext_factor_wide <-
bind_rows(ext_single, ext_multiple) %>%
as_tibble() %>%
select(Factor, Phenotype, STD_Genotype, STD_Genotype_SE, p_value, Beta) %>%
pivot_wider(id_cols=c(Phenotype, Factor),
            names_from=Beta, 
            values_from=c(STD_Genotype, STD_Genotype_SE, p_value))

ext_attenuation <-     
ext_factor_wide %>%
mutate(rg_Single=STD_Genotype_Single, rg_Multiple=STD_Genotype_Multiple,
       se_Single=as.numeric(STD_Genotype_SE_Single),
       se_Multiple=as.numeric(STD_Genotype_SE_Multiple)) %>%
mutate(attenuation_z=(rg_Single-rg_Multiple)/sqrt(se_Single^2+se_Multiple^2)) %>%
mutate(attenuation_p=2*pnorm(abs(attenuation_z), lower.tail=F)) %>%
select(Phenotype, Factor, p_value_Single, p_value_Multiple, attenuation_p)

knitr::kable(ext_attenuation)

```

